---
title: "Project Stage II"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

## Recent Grads
```{r}
recentgrads <- read_csv("Data/recentgrads.csv")

head(recentgrads)

recentgrads$recent_total <- recentgrads$Total
recentgrads$Total <- NULL

recentgrads$recent_men <- recentgrads$Men
recentgrads$Men <- NULL

recentgrads$recent_women <- recentgrads$Women
recentgrads$Women <- NULL

#recentgrads$recent_ShareWomen <- recentgrads$ShareWomen
#recentgrads$ShareWomen <- NULL

recentgrads$recent_sample_size <- recentgrads$Sample_size
recentgrads$Sample_size <- NULL

recentgrads$recent_employed <- recentgrads$Employed
recentgrads$Employed <- NULL

recentgrads$recent_full_time <- recentgrads$Full_time
recentgrads$Full_time <- NULL

recentgrads$recent_part_time <- recentgrads$Part_time
recentgrads$Part_time <- NULL

recentgrads$recent_full_time_year_round <- recentgrads$Full_time_year_round
recentgrads$Full_time_year_round <- NULL

recentgrads$recent_unemployed <- recentgrads$Unemployed
recentgrads$Unemployed <- NULL

recentgrads$recent_unemployment_rate <- recentgrads$Unemployment_rate
recentgrads$Unemployment_rate <- NULL

recentgrads$recent_median <- recentgrads$Median
recentgrads$Median <- NULL

recentgrads$recent_p25th <- recentgrads$P25th
recentgrads$P25th <- NULL

recentgrads$recent_p75th <- recentgrads$P75th
recentgrads$P75th <- NULL

recentgrads$recent_college_jobs <- recentgrads$College_jobs
recentgrads$College_jobs <- NULL

recentgrads$recent_non_college_jobs <- recentgrads$Non_college_jobs
recentgrads$Non_college_jobs <- NULL

recentgrads$recent_low_wage_jobs <- recentgrads$Low_wage_jobs
recentgrads$Low_wage_jobs <- NULL

head(recentgrads)
```


## All Grads
```{r}
allgrads <- read_csv("Data/allgrads.csv")




head(allgrads)

allgrads$all_total <- allgrads$Total
allgrads$Total <- NULL

allgrads$all_employed <- allgrads$Employed
allgrads$Employed <- NULL

allgrads$all_employed_full_time_year_round <- allgrads$Employed_full_time_year_round
allgrads$Employed_full_time_year_round <- NULL

allgrads$all_unemployed <- allgrads$Unemployed
allgrads$Unemployed <- NULL

allgrads$all_unemployment_rate <- allgrads$Unemployment_rate
allgrads$Unemployment_rate <- NULL

allgrads$all_median <- allgrads$Median
allgrads$Median <- NULL

allgrads$all_p25th <- allgrads$P25th
allgrads$P25th <- NULL

allgrads$all_p75th <- allgrads$P75th
allgrads$P75th <- NULL
```


## Grad students
```{r}
gradstudents <- read_csv("Data/gradstudents.csv")

allgrads <- read_csv("Data/allgrads.csv")


head(gradstudents)
```

```{r}
combinedrecentandgrad <- full_join(recentgrads, gradstudents, by = c("Major_code", "Major", "Major_category"))

totalcombination <- full_join(combinedrecentandgrad, allgrads, by = c("Major_code", "Major", "Major_category"))

totalcombination1 <- totalcombination %>% mutate(full_total = all_total + recent_total + Grad_total, full_employed = all_employed + recent_employed + Grad_total, total_unemployed = all_unemployed + recent_unemployed + Grad_unemployed)

totalcombination2 <- totalcombination1 %>% mutate(full_median = (recent_median + Grad_median + Nongrad_median + all_median)/4)

totalcombinationpaid <- totalcombination2 %>% arrange(desc(all_median))

write.csv(totalcombinationpaid, file = "totalcombinationpaid.csv")


topten <- filter(totalcombination2, all_median > 80000)
```

```{r}
ggplot(data = topten, aes(x = Major, y = all_median)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# PLOTTING ----------------

### Major Categories

```{r}
table(totalcombination2$Major_category)

totalcombination3 <- totalcombination2 %>% mutate(full_unemployment_rate = (recent_unemployment_rate + Grad_unemployment_rate + Nongrad_unemployment_rate + all_unemployment_rate)/4)

categories <- totalcombination3 %>% group_by(Major_category) %>% summarise(Count = n(), mean_income = mean(full_median), Total = sum(full_total), mean_unemployment = mean(full_unemployment_rate), mean_female = mean(ShareWomen)) %>% arrange(desc(Count))

#1
ggplot(data = categories, aes(x = fct_reorder(Major_category, Count, .desc = TRUE), y = Count)) + geom_bar(stat = "identity", fill = "blue") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Major Categories", x = "Major Category")

#2
ggplot(data = categories, aes(x = fct_reorder(Major_category, Total, .desc = TRUE), y = Total)) + geom_bar(stat = "identity", fill = "blue") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Total Cases", y = "Cases", x = "Major Category")
```
# 1. In our dataset, each unique major is classified as one of these 16 categories. Engineering is the most diverse category followed by; Education, Liberal Arts, Life Science, and Business.


### Major
```{r}
mostpopular <- totalcombination3 %>% arrange(desc(full_total)) %>% slice(1:25)

mostpopularcount <- mostpopular %>% group_by(Major_category) %>% summarise(Count = n()) %>% arrange(desc(Count))

#3
ggplot(data = mostpopular, aes(x = fct_reorder(Major, full_total, .desc = TRUE), y = full_total)) + geom_bar(stat = "identity", fill = "green") + theme(axis.text.x = element_text(angle = 75, hjust = 1)) + labs(title = "Popular Majors", x = "Majors", y = "Count")

#4
ggplot(data = mostpopularcount, aes(x = fct_reorder(Major_category, Count, .desc = TRUE), y = Count)) + geom_bar(stat = "identity", fill = "red") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Popular Major Categories", x = "Major Category")
```
#These are the 25 most popular majors in our dataset. Business Management and Administration is the most popular with Business being the most popular category amongst the top 25 majors.

```{r}
#5
ggplot(data = totalcombination2, aes(x = fct_reorder(Major, full_median, .desc = TRUE), y = full_median, color = Major_category)) + geom_point() + labs(main = "Median Income by Major", y = "Median Income", x = "Majors", color = "Major Category") + theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
# 4 This is the median income for all 173 majors we have in our data set. As you can see, the majors in the categories of Education, Engineering, and Health are the highest earners.

```{r}
#6
ggplot(data = categories, aes(x = fct_reorder(Major_category, mean_income, .desc = TRUE), y = mean_income)) + geom_bar(stat = "identity", fill = "blue") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Income per Major", x = "Major", y = "Median Income")
```
# These are the average median income for each major category in our dataset. As you can see engineering is paid the best. 




```{r}
#7
ggplot(data = mostpopular, aes(x = fct_reorder(Major, full_median, .desc = TRUE), y = full_median)) + geom_bar(stat = "identity", fill = "red") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Income per Major", x = "Major", y = "Median Income")

#8
ggplot(data = mostpopular, aes(x = fct_reorder(Major_category, full_median, .desc = TRUE), y = full_median)) + geom_boxplot(fill = "yellow") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Income for Popular Major Categories", x = "Major Category", y = "Median Income")
```
# 5 majors we want to investigate
- General Engineering 
- Economics
- Political Science and Government 
- Mathematics
- Biology
```{r}
ourmajors <- totalcombination3 %>% filter(Major %in% c("GENERAL ENGINEERING", "ECONOMICS", "POLITICAL SCIENCE AND GOVERNMENT", "MATHEMATICS", "BIOLOGY"))
```


#Unemployment
```{r}


#9
ggplot(data = totalcombination3, aes(x = fct_reorder(Major, full_unemployment_rate, .desc = TRUE), y = full_unemployment_rate, color = Major_category)) + geom_point() + labs(main = "Unemployment Rate by Major", y = "Unemployment Rate", x = "Majors", color = "Major Category") + theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

#10
ggplot(data = categories, aes(x = fct_reorder(Major_category, mean_unemployment, .desc = TRUE), y = mean_unemployment)) + geom_bar(stat = "identity", fill = "red") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Unemployment by Major Category", x = "Major Category", y = "Unemployment Rate")

#11
ggplot(data = mostpopular, aes(x = fct_reorder(Major, full_unemployment_rate, .desc = TRUE), y = full_unemployment_rate)) + geom_bar(stat = "identity", fill = "red") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "Unemployment by Major Category", x = "Major Category", y = "Unemployment Rate")
```

# Women
```{r}
#12
ggplot(data = totalcombination3, aes(x = fct_reorder(Major, ShareWomen, .desc = TRUE), y = ShareWomen, color = Major_category)) + geom_point() + labs(main = "% Female", y = "% Female", x = "Majors", color = "Major Category") + theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

#13
ggplot(data = categories, aes(x = fct_reorder(Major_category, mean_female, .desc = TRUE), y = mean_female)) + geom_bar(stat = "identity", fill = "pink") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "% Female by Major Category", x = "Major Category", y = "% Female")

#14
ggplot(data = mostpopular, aes(x = fct_reorder(Major, ShareWomen, .desc = TRUE), y = ShareWomen)) + geom_bar(stat = "identity", fill = "red") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(title = "% Female by Major Category", x = "Major", y = "% Female")
```

# Closer look at our majors 
```{r}
#15 
ggplot(data = ourmajors, aes(x = Major, y = full_median, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#16
ggplot(data = ourmajors, aes(x = Major, y = full_unemployment_rate, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#17
ggplot(data = ourmajors, aes(x = Major, y = ShareWomen, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#18
ggplot(data = ourmajors, aes(x = Major, y = full_median, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1))

#19
nongrad <- ggplot(data = ourmajors, aes(x = Major, y = Nongrad_median, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.ticks.x=element_blank())

grad <- ggplot(data = ourmajors, aes(x = Major, y = Grad_median, fill = Major_code)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 60, hjust = 1), axis.ticks.x=element_blank())

grid.arrange(grad, nongrad)
```










### Plots
```{r}
outliers <- totalcombination3 %>%
  filter(full_total >= 310000 | full_median >= 65000)

ggplot(totalcombination3, aes(Major, full_total)) +
  geom_point() + 
  ggrepel::geom_label_repel(aes(label=paste("(",Major,",",full_median,")"), vjust = 4, hjust = 0, colour = Major_category),
    data = outliers, show.legend = FALSE)
```


```{r}
outliers1 <- totalcombination1 %>%
  filter(ShareWomen <= 0 | ShareWomen >= 	0.9679982)

ggplot(totalcombination1, aes(Major, full_total, fill = recent_men)) +
  geom_col() +
  ggrepel::geom_label_repel(aes(label=paste("(",Major,",",recent_men," "," ",recent_low_wage_jobs,")"), vjust = 100, colour = Major_category),
    data = outliers1, show.legend = FALSE)
```

```{r}
outliers2 <- totalcombination1 %>%
  filter(recent_unemployment_rate >= .14 | recent_low_wage_jobs >= 25000)

ggplot(totalcombination1, aes(Major, recent_unemployment_rate)) +
  geom_jitter() +
  ggrepel::geom_label_repel(aes(label=paste("(",Major,",",recent_low_wage_jobs,")"), vjust = 4, hjust = 0, colour = Major_category),
    data = outliers2, show.legend = FALSE)
```
